Boosting Higgs Pair Production in the $b\bar{b}b\bar{b}$ Final State with Multivariate Techniques

### 2015-12-14 paper/sec-mva.tex
completed Daniela's comments

In this section, first of all we present the specific MVA that we use,
based on feed-forward multi-layer neural networks.
%
We then introduce the input variables that are
used the MVA, including the jet substructure
variables, and then present the results of $S/\sqrt{B}$ including
the MVA effects.
%
Since we find that jet substructure variables are a crucial input
to the MVA, we show representative kinematic distributions
for these variables.
%
Finally, we assess the robustness of the MVA strategy in
high-PU environments.

@@

It important to verify, for each value of
the cut in the ANN output $y_{\rm cut}$, how many
signal and background events are expected at the HL-LHC,
with an integrated luminosity of $\mathcal{L}=3$ ab$^{-1}$.
%
This comparison is shown in 
Fig.~\ref{fig:nev2}.
%
As can be seen,
in the boosted category, for a value $y_{\rm cut}\simeq 0.8$
we end up with 800 signal events and around $2\cdot 10^4$ background
events.
%
Similar results are obtained in the intermediate and resolved
categories: in the former we find 700 ($6\cdot 10^3$) signal (background)
events for $y_{\rm cut}\simeq 0.8$, and in the latter
1300 ($8\cdot 10^4$) signal (background) events for
$y_{\rm cut}\simeq 0.65$.
%
Note that the MVA achieves a very substantial background suppression
with only a
moderate reduction of signal efficiency in exchange.

A useful property of MVAs such as the one used in this work
is that they provide direct  physical insight about which of the
input variables contribute the most to the separation between
signal and background.
%
In the case of ANNs, this can be quantified by computing the sum
of the absolute value of all the weights connected to a given
input neuron $i$, that is
\be
\label{eq:totweight}
\omega^{\rm (tot)}_i \equiv \sum_{k=1}^{n^{(2)}} \Big|\omega^{(2)}_{ki}\Big| \, ,
\qquad i=1,\ldots,N_{\rm var} \, ,
\ee
with $\omega^{(2)}_{ki}$ the value of the weight connecting
the $k$-th neutron of the second layer with the $i$-th neuron of
the first (input) layer, and $n^{(2)}=5$ the number of
neurons in the second layer.
%
Those input variables with a larger value of $\omega^{\rm (tot)}_i$ will be those
that play a dominant role in enhancing the signal
discrimination using the MVA.

@@

The results for the signal significance $S/\sqrt{B}$ and
the signal over background ratio
$S/B$ as a function of $y_{\rm cut}$
for the three categories are
Fig.~\ref{fig:sb_mva}.
%
The values 
for $y_{\rm cut}=0$ correspond to those at
the end of the cut-based analysis.
%
We observe how in the three
 categories there is a marked  improvement in signal
significance as compared to the pre-MVA results.
%
We also observe a substantial enhancement in $S/B$, arising
from the background suppression indeed by the MVA, reaching
values of 1\%, 6\% and 3.5\% in the resolved,
intermediate and boosted categories.
%
This improvement in $S/B$ is crucial to ensure the feasibility
of this measurement, since it allows systematic
uncertainties in the background determination to
be at the same level of a few percent level.

@@

From Table~\ref{table:cutflowMVA} we see that at the end of the
MVA the signal significance in the boosted category increases
from 1.2 to 5.4, with $S/B$ increasing from $0.14\%$ to $3.4\%$,
with almost 800 signal events expected at the HL-LHC.
%
For the intermediate and resolved category, $S/\sqrt{B}$
increases from 0.9 and 1.0 to 4.8 and 4.5, with
the signal over background ratio raising from
$0.12\%$ and $0.03\%$ to 6\% and 1\$, respectively.
%
The overall combined significance is thus $S\sqrt{B}\simeq 8$,
well above the threshold for discovery.
%
These remarkable results are however subject to a justified
criticism:
the HL-LHC will be a high-PU environment,
which will affect the description of the various
kinematical distributions used as input to the MVA.
%
Therefore, next we quantify the robustness of the
results summarized Table~\ref{fig:sb_mva}
in a realistic high-PU environment.

@@

Another important result from Table~\ref{table:cutflowMVA_PU} is that,
for the optimal MVA cut, the signal over background ratio
is not unfeasibly small: we obtain that $S/B$ is 3.4\% (1.5\%)
in the boosted (resolved) categories.
%
This indicates that while this measurement is still highly challenging,
requiring a careful extraction of the QCD
background from the data, it is certainty within reach.
%
As in the case without PU, 
our analysis  points out that both
$S/B$ and $S/\sqrt{B}$ could be increased  by reducing the background component
that arises from the mis-identification of light and charm
jets as $b$-jets.

@@

In Fig.~\ref{fig:sb_mva_PU} we show the signal significance,
$S/\sqrt{B}$, and the signal over background ratio,
$S/B$, accounting now for the effects of PU.
%
The corresponding results in the case without PU were shown in
Fig.~\ref{fig:sb_mva}.
%
As can be seen, the MVA-driven enhancement is robust in the
presence of PU.
%
While as expected there is some degradation as compared to the case
without PU,
it is still possible to
achieve a signal significance of
around $S/\sqrt{B}$ between 2.5 and 3.0 (depending on the
specific choice of $y_{\rm cut}$)
separately for the boosted and resolved
categories.
%
On the other hand, the intermediate category now becomes marginal.
%
We conclude that the qualitative conclusions drawn
in the case without PU also hold when the analysis
is performed in a high-PU environment.
%
Since no specific effort has been performed to
optimize PU subtraction, for instance by tuning the value
of the patch length $a$ in {\tt SoftKiller}, we believe that
there could be still room for further  improvement.

@@

It is useful to quantify which of the input variables
to the MVA carry the highest discrimination power
in the case of PU,
and compare this with the corresponding
results without PU, by means of
Eq.~(\ref{eq:totweight}).
%
The 
results for the resolved and boosted categories are shown
on Fig.~\ref{fig:nnweights_PU}.
%
As we can see by comparing with Fig.~\ref{fig:nnweights}, 
in the boosted category, the discrimination power of the invariant
mass of Higgs candidates is decreased, while
that of  substructure
variables such as $C_2^{(\beta)}$ and
$D^{(\beta)}_2$ is conversely
increased.
%
In addition, also the $p_T$ distributions of the AKT03
subjets within the large-$R$
jets carry important information.
%
In the case of the resolved category,  the highest
values of the ANN weights without PU
were found for the $p_T$ of the 
Higgs candidates and their invariant masses, as well
as the for invariant mass of the di-Higgs system
and the $p_T$ of the small-$R$ jets.
%
As we can see in Fig.~\ref{fig:nnweights_PU},
these results  are reasonably stable in the case with PU.

@@

We can finally combine the three categories, resolved,
intermediate and boosted, and determine the overall signal
significance, finding
\be
\lp \frac{S}{\sqrt{B}}\rp_{\rm tot} \simeq 4.0 \, ,\quad
\mathcal{L}=3\,{\rm ab}^{-1}\, .
\ee
%
We thus conclude that the measurement of
Higgs pair production in the $b\bar{b}b\bar{b}$ final state
should be 
well above the threshold for evidence, and even meeting the
requirements for discovery might be within reach, provided it is possible
to suppress the background from light and charm jet mis-identifications.
%
Extrapolating the above results
to total integrated luminosity expected at the end of
LHC Run II, one finds
instead
\be
\lp \frac{S}{\sqrt{B}}\rp_{\rm tot} \simeq 1.3(2.6) \, ,\quad
\mathcal{L}=300\,{\rm fb}^{-1}\, ,
\ee
where the number in brackets corresponds to neglecting all backgrounds
except for the QCD $4b$ multijets.
%
This suggest that, remarkably, accessing Higgs pair production at Run II
might not be completely impossible, specially accounting
for the fact that the mean number of PU events per
bunch crossing is much smaller
that at the HL-LHC.

>>>>

In this section, first of all we present the specific MVA that we use,
based on feed-forward multi-layer neural networks.
%
We then introduce the input variables that are
used in the MVA, including the jet substructure
variables, and then present the signal significance obtained
by applying the MVA.
%
Then we assess the robustness of the MVA strategy in
high-PU environments.

@@

It important to verify, for each value of
the cut in the ANN output $y_{\rm cut}$, how many
signal and background events are expected at the HL-LHC,
with an integrated luminosity of $\mathcal{L}=3$ ab$^{-1}$.
%
This comparison is shown in 
Fig.~\ref{fig:nev2}.
%
We observe that
in the boosted category, for a value $y_{\rm cut}\simeq 0.8$
we end up with 800 signal events and around $2\cdot 10^4$ background
events.
%
Similar results are obtained in the intermediate and resolved
categories: in the former we find 700 ($6\cdot 10^3$) signal (background)
events for $y_{\rm cut}\simeq 0.8$, and in the latter
1300 ($8\cdot 10^4$) signal (background) events for
$y_{\rm cut}\simeq 0.65$.
%
Note that the MVA achieves a very substantial background suppression
with only a
moderate reduction of signal efficiency.

A useful property of MVAs, as the one used in our
analysis,
is that they provide direct  physical insight about which of the
input variables contribute the most to the separation between
signal and background.
%
In the case of ANNs, this can be quantified by computing the sum
of the absolute value of all the weights connected to a given
input neuron $i$, that is
\be
\label{eq:totweight}
\omega^{\rm (tot)}_i \equiv \sum_{k=1}^{n^{(2)}} \Big|\omega^{(2)}_{ki}\Big| \, ,
\qquad i=1,\ldots,N_{\rm var} \, ,
\ee
with $\omega^{(2)}_{ki}$ the value of the weight connecting
the $k$-th neutron of the second layer with the $i$-th neuron of
the first (input) layer, and $n^{(2)}=5$ the number of
neurons in the second layer.
%
Those input variables with a larger value of $\omega^{\rm (tot)}_i$ will be those
that play a dominant role in enhancing the signal
discrimination using the MVA.

@@

The results for the signal significance $S/\sqrt{B}$ and
the signal over background ratio
$S/B$ as a function of $y_{\rm cut}$
for the three categories are
Fig.~\ref{fig:sb_mva}.
%
The values 
for $y_{\rm cut}=0$ correspond to those at
the end of the cut-based analysis.
%
We observe how in the three
 categories there is a marked  improvement in signal
significance as compared to the pre-MVA results.
%
We also observe a substantial enhancement in $S/B$, arising
from the background suppression achieved by the MVA, reaching
values of 1\%, 6\% and 3.5\% in the resolved,
intermediate and boosted categories.
%
This improvement in $S/B$ is crucial to ensure the feasibility
of this measurement, since it allows systematic
uncertainties in the background determination to
be at the same level of a few percent level.

@@

From Table~\ref{table:cutflowMVA} we see that
after we apply the MVA
the signal significance in the boosted category increases
from 1.2 to 5.4, with $S/B$ increasing from $0.14\%$ to $3.4\%$,
with almost 800 signal events expected at the HL-LHC.
%
For the intermediate and resolved category, $S/\sqrt{B}$
increases from 0.9 and 1.0 to 4.8 and 4.5, with
the signal over background ratio raising from
$0.12\%$ and $0.03\%$ to 6\% and 1\$, respectively.
%
The overall combined significance is thus $S\sqrt{B}\simeq 8$,
well above the threshold for discovery.
%
These remarkable results are however subject to a justified
criticism:
the HL-LHC will be a high-PU environment,
which will affect the description of the various
kinematical distributions used as input to the MVA.
%
Therefore, next we quantify the robustness of the
results summarized Table~\ref{fig:sb_mva}
in a realistic high-PU environment.

@@

Another important result from Table~\ref{table:cutflowMVA_PU} is that,
for the optimal MVA cut, the signal over background ratio
is 3.4\% (1.5\%)
in the boosted (resolved) categories.
%
This indicates that while this measurement is still highly challenging,
requiring a careful extraction of the QCD
background from the data, it is certainty within reach.
%
As in the case without PU, 
our analysis  points out that both
$S/B$ and $S/\sqrt{B}$ could be increased  by reducing the background component
that arises from the mis-identification of light and charm
jets as $b$-jets.

@@

In Fig.~\ref{fig:sb_mva_PU} we show the signal significance,
$S/\sqrt{B}$, and the signal over background ratio,
$S/B$, accounting now for the effects of PU.
%
The corresponding results in the case without PU were shown in
Fig.~\ref{fig:sb_mva}.
%
As can be seen, the MVA-driven enhancement is robust in the
presence of PU.
%
While
there is some degradation as compared to the case
without PU,
it is still possible to
achieve a signal significance of
around $S/\sqrt{B}$ between 2.5 and 3.0 (depending on the
specific choice of $y_{\rm cut}$)
separately for the boosted and resolved
categories.
%
On the other hand, the intermediate category is now marginal.
%
We conclude that the qualitative conclusions drawn
in the case without PU also hold when the analysis
is performed in a high-PU environment.
%
Since no specific effort has been performed to
optimize PU subtraction, for instance by tuning the value
of the patch length $a$ in {\tt SoftKiller}, we believe that
there could be still room for further  improvement.

@@

It is useful to quantify which of the input variables
to the MVA carry the highest discrimination power
in the case of PU,
and compare this with the corresponding
results without PU, by means of
Eq.~(\ref{eq:totweight}).
%
The 
results for the resolved and boosted categories are shown
on Fig.~\ref{fig:nnweights_PU}.
%
As we can see by comparing with Fig.~\ref{fig:nnweights}, 
in the boosted category, the discrimination power of the invariant
mass of Higgs candidates is decreased, while
that of  substructure
variables such as $C_2^{(\beta)}$ and
$D^{(\beta)}_2$ is conversely
increased.
%
In addition, also the $p_T$ distributions of the AKT03
subjets within the large-$R$
jets carry important information.
%
In the case of the resolved category,  the highest
values of the ANN weights without PU
were found for the $p_T$ of the 
Higgs candidates and their invariant masses, as well
as the for invariant mass of the di-Higgs system
and the $p_T$ of the small-$R$ jets.
%
As we can see in Fig.~\ref{fig:nnweights_PU},
this is the case also with PU.

@@

We can finally combine the three categories, resolved,
intermediate and boosted, and determine the overall signal
significance, finding
\be
\lp \frac{S}{\sqrt{B}}\rp_{\rm tot} \simeq 4.0 \, ,\quad
\mathcal{L}=3\,{\rm ab}^{-1}\, .
\ee
%
We thus conclude that the measurement of
Higgs pair production in the $b\bar{b}b\bar{b}$ final state
should be 
well above the threshold for evidence.
%
Even
discovery might be within reach if the
misidentification of light and charm jets as b-jets can be  reduced.
%
Extrapolating the above results
to total integrated luminosity expected at the end of
LHC Run II, one finds
instead
\be
\lp \frac{S}{\sqrt{B}}\rp_{\rm tot} \simeq 1.3(2.6) \, ,\quad
\mathcal{L}=300\,{\rm fb}^{-1}\, ,
\ee
where the number in brackets corresponds to neglecting all backgrounds
except for the QCD $4b$ multijets.
%
This suggest that, remarkably, accessing Higgs pair production at Run II
might not be completely impossible, specially accounting
for the fact that the mean number of PU events per
bunch crossing is much smaller
that at the HL-LHC.

### 2015-12-14 paper/sec-results.tex
more work on the text
completed Daniela's comments

In this section we present the results of the 
cut-based analysis and provide a
cut-flow for the different analysis steps.
%
We
compare our results with other recent studies,
and assess how the signal significance
is affected if only the $4b$ component of the
QCD multi-jet background is taken into account but not 
the $2b2j$ and $4j$ components.
%
We present now the results in the no PU scenario, the impact
of PU will be discussed in the next section.

@@

In Table~\ref{tab:cutflowdetails}
we summarize the definition of the different
levels of the cut-flow of the present analysis,
separated into the boosted, intermediate,
    and resolved categories.
    %
    From top to bottom, additional cuts are required in addition
    to the previous ones.
    %
    In more details, these different analysis levels are the following:

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}[t]
  \centering
  \begin{tabular}{|c|c|c|c|}
\hline
&  Boosted  &   Intermediate &  Resolved  \\
\hline
\hline
{\bf C0} &  \multicolumn{3}{c|}{Generator level} \\
\hline
{\bf C1a} & $N_{\rm jets}^{R10}\ge 2$ & $N_{\rm jets}^{R04}\ge 2$, $N_{\rm jets}^{R10}=1$  &
$N_{\rm jets}^{R04}\ge 4$ \\
\hline
{\bf  C1b} & \multicolumn{3}{c|}{+$p_T$ cuts} \\
{\bf C1c} & \multicolumn{3}{c|}{+rapidity cuts}\\
\hline
 {\bf C1d} & +$N_{\rm MDT}\ge 2$ & +$N_{\rm jets}^{R10}=1$ with MDT  &
 +Higgs reconstruction \\
 \hline
{\bf C1e} & \multicolumn{3}{c|}{ +$m_H$ window cut} \\
\hline
{\bf C2} & \multicolumn{3}{c|}{+$b$-tagging}    \\
\hline
  \end{tabular}
  \caption{\small Definition of the cut-flow levels for the boosted, intermediate
    and resolved selections.
    %
      \label{tab:cutflowdetails}
  }
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

    \begin{itemize}
    \item {\bf C0}:  generator level cross-sections, with
      mild generation cuts in the case of the background processes, and no
      generation cuts
      for the signal events.
      %
    \item {\bf C1a}:  check that we have at least
      two large-$R$ jets (in the boosted case),
      one large-$R$ jet and at least 2 small-$R$ jets (in the intermediate
      case) and at least four small-$R$ jets (in the resolved case).
      %
      No acceptance requirements are imposed.
    \item {\bf C1b}: require that the above jets should
      satisfy the corresponding $p_T$ thresholds,
      $p_T \ge 200$ GeV for large-$R$ jets and
      $p_T \ge 50$ GeV for small-$R$ jets.
    \item {\bf C1c}: same for the
      rapidity acceptance requirement.
    \item {\bf C1d}: the two leading large-$R$ jets must
      be mass-drop tagged in the boosted category.
      %
      In the intermediate category, the large-$R$ jet must also be mass-drop tagged.
      %
    \item {\bf C1e}: after the two Higgs candidates  have been reconstructed,
      their invariant masses are required to lie a window around $m_H$,
      in particular between 85 and 165 GeV.
          \item {\bf C2}: the
            $b$-tagging conditions are
            imposed, see
            Sect.~\ref{sec:btagging}.
      \end{itemize}
    Signal and background events satisfying all the analysis cut up to the
    C2 level
    are then used as input for the MVA training, to be described next
    in Sect.~\ref{sec:mva}.
    %
    In Table~\ref{tab:cutflow_noPU_1} we collect
    the values for cross-sections for signal and background processes
    for different levels of the cut flow.
    %
    Results are divided into the resolved, intermediate and boosted categories,
    and are inclusive up to the C2 level, where exclusivity is imposed.
    %

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}[t]
  \centering
  \scriptsize
  \input{tables/table_noPU_1.tex}
  $\,$ \\
  \vspace{0.5cm}
  \input{tables/table_noPU_2.tex}
  $\,$ \\
  \vspace{0.5cm}
    \input{tables/table_noPU_3.tex}
    \caption{\small The cross-sections, in femtobarns,
      for the signal and the various background
      processes at different steps of the
      cut-flow, for the resolved (upper table),
      intermediate (middle table) and boosted
      (lower table) categories, for the analysis
      without PU.
      %
      In each case, we also provide the signal over
      background ratio, $S/B$, and the signal
      significance, $S/\sqrt{B}$, considering either
      the total background or only the $4b$ component, for
      a total integrated luminosity of $\mathcal{L}=3$ ab$^{-1}$.
      %
      The different levels of the cut-flow are summarized
      in Table~\ref{tab:cutflowdetails}.
 \label{tab:cutflow_noPU_1}}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

 It is interesting to compare in each category the interplay
 between the reducible and irreducible components of the
 QCD backgrounds.
 %
 In all cases, the $4b$ and $2b2j$ have comparable
 magnitudes, taking into account the size that the higher-order QCD
 effects could induce in
 the corresponding normalizations.
 %
 On the other hand, the $4j$ component
is always smaller by at least
 one order of magnitude.
 %
  So while the $4j$ component can be safely
 neglected, the inclusion of the
 $2b2j$ component is essential to robustly assess the feasibility
 of measuring Higgs pairs in this final state, and this is markedly
 important
 in the boosted category.
 %
 This has the important
 consequence that a promising avenue to improve the prospects
 of this measurement would be to reduce, as much as possible,
 the light and charm jet mis-identification rate.

@@

The reason why the  $2b2j$ process
cannot be neglected as compared to the $4b$ component
is the following.
%
In the boosted category, for example,
the cross-section  before
$b$-tagging is more than two orders
of magnitude larger in the $2b2j$
sample as compared to the $4b$ sample.
%
Now, after $b$-tagging, the naive expectation would
be a suppression of the former by a factor $(f_l/f_b)^2 \simeq
1.5\cdot 10^{-4}$, as compared to the $4b$ component,
since a total of four $b$-tags are required to classify the
event as a Higgs candidate.
%
So the ratio of $2b2j$ over $4b$ should be
around $\simeq 3\%$, and therefore negligible.
  %
While  we have checked that this expectation is borne
out at the parton level,
we find that  when parton shower effects
are accounted for, due both to radiation of $b\bar{b}$ pairs
and from selection effects, the situation is different.
%
In this case,
the
number of  $b$ quarks in the  final state is
increased substantially in the $2b2j$ component as compared
to the parton level content, while at the same
time the number of events in the $4b$ sample
with 4 $b$-jets is reduced.

We can make these statements more quantitative in the following way.
%
To first approximation, and neglecting the contribution from
charm mis-identification,
the
overall efficiency of the $b$-tagging requirements in the resolved category will be
given by the following expression:
\be
\label{btaggingeff}
{\rm EFF}_{\rm b-tag}\simeq \sum_{j=0}^{4}n^{\rm (b-jet)}_j\cdot f_b^{j}\cdot f_l^{4-j} \, ,
\ee
with $n^{\rm (b-jet)}_j$ the fraction of events, satisfying all the selection cuts,
where $j$ jets, out of the leading four jets of the event,
contain $b$ quarks (with $p_T^b\ge 15$
GeV).
%
Similar expressions can be derived to the boosted and intermediate categories.
%
The naive expectation is that all events in the $4b$ sample have $n^{\rm (b-jet)}_4\simeq 1$
and $n^{\rm (b-jet)}_j\simeq 0$ for $j\ne 4$, while the events in the $2b2j$ sample
should have $n^{\rm (b-jet)}_2\simeq 1$ and zero otherwise.
%
This leads to a ratio of overall $b$-tagging selection efficiencies
\be
\label{eq:naive}
\frac{ {\rm EFF}_{\rm b-tag} \lc 2b2j \rc}{{\rm EFF}_{\rm b-tag} \lc 4b\rc}
  \simeq
 \lp \frac{f_l}{f_b}\rp^2 \simeq 1.5\cdot 10^{-4} \, .
\ee
However, after the parton shower the above is not a good approximation.
%
First of all, we will have a non-negligible fractions $n^{\rm (b-jet)}_j$
with $j=3,4$ also in the $2b2j$ sample, due to $b$-quark pair radiation
during the shower.
%
Secondly, not all events in the $4b$ sample will lead to four small-$R$ $b$-jets,
due to a combination of selection cuts and
parton shower effects.
%

@@

In Table~\ref{tab:btaggingcheck} we collect
the values of $n^{\rm (b-jet)}_j$ for the signal and the three QCD background samples.
%
We find that, rather than the estimate Eq.~(\ref{eq:naive}),
the correct ratio of $b$-tagging selection efficiencies is instead
\be
\frac{{\rm EFF}_{\rm b-tag} \lc 2b2j\rc}{{\rm EFF}_{\rm b-tag} \lc 4b\rc}=
  \frac{0.04\%}{8.4\%} \simeq 5\cdot 10^{-3} \, .
  \ee
  This suppression factor is of the same order of the ratio of $4b$ to $2b2j$ cross-sections
  in the resolved category before $b$-tagging, see Table~\ref{tab:cutflow_noPU_1}.
    %
    This explains why the $2b2j$ contribution cannot be neglected as compared
    to the irreducible $4b$ component of the QCD background.
    %
    A similar calculation from the numbers of Table~\ref{tab:btaggingcheck} shows
    that, on the other hand, the $4j$ component of the background can safely
    be neglected.

@@

%

>>>>

In this section we present the results of the 
cut-based analysis and provide a
cut-flow for the different analysis steps.
%
We
compare our results with other recent studies,
and assess how the signal significance
is affected if only the $4b$ component of the
QCD multi-jet background is taken into account but not 
the $2b2j$ and $4j$ components.
%
First we present the results without considering PU,
followed by the corresponding analysis of the PU effect.

@@

In Table~\ref{tab:cutflowdetails}
we summarize the definition of the different
steps of the cut-flow of the present analysis,
separated into the boosted, intermediate,
    and resolved categories.
    %
    From top to bottom, additional cuts are required in addition
    to the previous ones.
    %
    In more details, these different analysis steps are the following:

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}[t]
  \centering
  \begin{tabular}{|c|c|c|c|}
\hline
&  Boosted  &   Intermediate &  Resolved  \\
\hline
\hline
{\bf C0} &  \multicolumn{3}{c|}{Generator level} \\
\hline
{\bf C1a} & $N_{\rm jets}^{R10}\ge 2$ & $N_{\rm jets}^{R04}\ge 2$, $N_{\rm jets}^{R10}=1$  &
$N_{\rm jets}^{R04}\ge 4$ \\
\hline
{\bf  C1b} & \multicolumn{3}{c|}{+$p_T$ cuts} \\
{\bf C1c} & \multicolumn{3}{c|}{+rapidity cuts}\\
\hline
 {\bf C1d} & +$N_{\rm MDT}\ge 2$ & +$N_{\rm jets}^{R10}=1$ with MDT  &
 +Higgs reconstruction \\
 \hline
{\bf C1e} & \multicolumn{3}{c|}{ +$m_H$ window cut} \\
\hline
{\bf C2} & \multicolumn{3}{c|}{+$b$-tagging}    \\
\hline
  \end{tabular}
  \caption{\small Definition of the cut-flow steps for the boosted, intermediate
    and resolved selections.
    %
      \label{tab:cutflowdetails}
  }
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

    \begin{itemize}
    \item {\bf C0}:  generator level cross-sections, with
      mild generation cuts in the case of the background processes, and no
      generation cuts
      for the signal events.
      %
    \item {\bf C1a}:  check that we have at least
      two large-$R$ jets (in the boosted case),
      one large-$R$ jet and at least 2 small-$R$ jets (in the intermediate
      case) and at least four small-$R$ jets (in the resolved case).
      %
      No acceptance requirements are imposed.
    \item {\bf C1b}: require that the above jets should
      satisfy the corresponding $p_T$ thresholds,
      $p_T \ge 200$ GeV for large-$R$ jets and
      $p_T \ge 50$ GeV for small-$R$ jets.
    \item {\bf C1c}: same for the
      rapidity acceptance requirement.
    \item {\bf C1d}: the two leading large-$R$ jets must
      be mass-drop tagged in the boosted category.
      %
      In the intermediate category, the large-$R$ jet must also be mass-drop tagged.
      %
    \item {\bf C1e}: after the two Higgs candidates  have been reconstructed,
      their invariant masses are required to lie a window around $m_H$,
      in particular between 85 and 165 GeV.
          \item {\bf C2}: the
            $b$-tagging conditions are
            imposed, see
            Sect.~\ref{sec:btagging}.
      \end{itemize}
    Signal and background events satisfying all the analysis cut up to the
    C2 level
    are then used as input for the MVA training, to be described next
    in Sect.~\ref{sec:mva}.
    %
    In Table~\ref{tab:cutflow_noPU_1} we collect
    the values for cross-sections for signal and background processes
    for different steps of the cut flow.
    %
    Results are divided into the resolved, intermediate and boosted categories,
    and are inclusive up to the C2 level, where exclusivity is imposed.
    %

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}[t]
  \centering
  \scriptsize
  \input{table_noPU_1.tex}
  $\,$ \\
  \vspace{0.5cm}
  \input{table_noPU_2.tex}
  $\,$ \\
  \vspace{0.5cm}
    \input{table_noPU_3.tex}
    \caption{\small The cross-sections, in femtobarns,
      for the signal and the various background
      processes at different steps of the
      cut-flow, for the resolved (upper table),
      intermediate (middle table) and boosted
      (lower table) categories, for the analysis
      without PU.
      %
      In each case, we also provide the signal over
      background ratio, $S/B$, and the signal
      significance, $S/\sqrt{B}$, considering either
      the total background or only the $4b$ component, for
      a total integrated luminosity of $\mathcal{L}=3$ ab$^{-1}$.
      %
      The different steps of the cut-flow are summarized
      in Table~\ref{tab:cutflowdetails}.
 \label{tab:cutflow_noPU_1}}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

 It is interesting to compare in each category the interplay
 between the reducible and irreducible components of the
 QCD backgrounds.
 %
 In all cases, the $4b$ and $2b2j$ have comparable
 magnitudes, taking into account the size that the higher-order QCD
 effects could induce in
 the corresponding normalizations.
 %
 On the other hand, the $4j$ component
is always smaller by at least
 one order of magnitude.
 %
  So while the $4j$ component can be safely
 neglected, the inclusion of the
 $2b2j$ component is essential to robustly assess the feasibility
 of measuring Higgs pairs in this final state,
 specially
 in the boosted category.
 %
 This has the important
 consequence that a promising avenue to improve the prospects
 of this measurement would be to reduce, as much as possible,
 the light and charm jet mis-identification rate.

@@

The reason why the  $2b2j$ process
cannot be neglected as compared to the $4b$ component
is the following.
%
In the boosted category, for example,
the cross-section  before
$b$-tagging is more than two orders
of magnitude larger in the $2b2j$
sample as compared to the $4b$ sample.
%
Now, after $b$-tagging, the naive expectation would
be a suppression of the former by a factor $(f_l/f_b)^2 \simeq
1.5\cdot 10^{-4}$, as compared to the $4b$ component,
since a total of four $b$-tags are required to classify the
event as a Higgs candidate.
%
So the ratio of $2b2j$ over $4b$ should be
around $\simeq 3\%$, and therefore negligible.
  %
While  we have checked that this expectation is borne
out at the parton level,
we find that  when parton shower effects
are accounted for, due both to radiation of $b\bar{b}$ pairs
and from selection effects, the situation is different.
%
In this case,
the
number of  $b$ quarks in the  final state is
increased substantially in the $2b2j$ component as compared
to the parton level, while at the same
time the number of events in the $4b$ sample
with 4 $b$-jets is reduced.

We can make these statements more quantitative in the following way.
%
To first approximation, and neglecting the contribution from
charm mis-identification,
the
overall efficiency of the $b$-tagging requirements in the resolved category will be
given by the following expression:
\be
\label{btaggingeff}
{\rm EFF}_{\rm b-tag}\simeq \sum_{j=0}^{4}n^{\rm (b-jet)}_j\cdot f_b^{j}\cdot f_l^{4-j} \, ,
\ee
with $n^{\rm (b-jet)}_j$ the fraction of events, satisfying all the selection cuts,
where $j$ jets, out of the leading four jets of the event,
contain $b$ quarks (with $p_T^b\ge 15$
GeV).
%
Similar expressions can be derived for
the boosted and intermediate categories.
%
The naive expectation is that all events in the $4b$ sample have $n^{\rm (b-jet)}_4\simeq 1$
and $n^{\rm (b-jet)}_j\simeq 0$ for $j\ne 4$, while the events in the $2b2j$ sample
should have $n^{\rm (b-jet)}_2\simeq 1$ and zero otherwise.
%
This leads to a ratio of overall $b$-tagging selection efficiencies
\be
\label{eq:naive}
\frac{ {\rm EFF}_{\rm b-tag} \lc 2b2j \rc}{{\rm EFF}_{\rm b-tag} \lc 4b\rc}
  \simeq
 \lp \frac{f_l}{f_b}\rp^2 \simeq 1.5\cdot 10^{-4} \, .
\ee
However, after the parton shower the above is not a good approximation.
%
First of all, we will have a non-negligible fractions $n^{\rm (b-jet)}_j$
with $j=3,4$ also in the $2b2j$ sample, due to $b$-quark pair radiation
during the shower.
%
Secondly, not all events in the $4b$ sample will lead to four small-$R$ $b$-jets,
due to a combination of selection cuts and
parton shower effects.
%

@@

In Table~\ref{tab:btaggingcheck} we collect
the values of $n^{\rm (b-jet)}_j$ for the signal and the three QCD background samples.
%
We find that, rather than the estimate Eq.~(\ref{eq:naive}),
the correct ratio of $b$-tagging selection efficiencies is instead
\be
\frac{{\rm EFF}_{\rm b-tag} \lc 2b2j\rc}{{\rm EFF}_{\rm b-tag} \lc 4b\rc}=
  \frac{0.04\%}{8.4\%} \simeq 5\cdot 10^{-3} \, .
  \ee
  This suppression factor is of the same order of the ratio of $4b$ to $2b2j$ cross-sections
  in the resolved category before $b$-tagging, shown
  in Table~\ref{tab:cutflow_noPU_1}.
    %
    This explains why the $2b2j$ contribution cannot be neglected as compared
    to the irreducible $4b$ component of the QCD background.
    %
    A similar calculation from the numbers in Table~\ref{tab:btaggingcheck} shows
    that, on the other hand, the $4j$ component of the background can safely
    be neglected.

@@
